---
title: "Final Project"
author: "AFN"
date: "2026-04-17"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tibble)
library(dplyr)
library(tseries)
library(fpp2)
library(ggplot2)
library(urca)
library(TSA)
library(MTS)
library(flextable)
```

```{r read in data}
traffic <- read.csv('Data/daily-website-visitors.csv')
```

This project uses data recording internet traffic at various websites. The variables of interest include the number of page loads each day, the number of unique visitors each day, and the number of returning visitors each day. We will first perform univariate time series analysis on each of the three variables to forecast the next X observations. Next, we will perform multivariate time series analysis using all three variables to forecast the next X observations. Finally, we will perform dynamic regression using returning visitors as the outcome variable and page loads and unique visitors as predictors. 

# Univariate Unique Visits 

## Daily Page Loads

```{r}
web_ts <- ts(traffic %>%
               select(Page.Loads),
             start = c(1,1),
             frequency = 7)
pred_num <- round(length(web_ts)*0.1, 0)
train_num <- length(web_ts) - pred_num
web_train <- head(web_ts, train_num)
web_test <- tail(web_ts, pred_num)
```

```{r, dpi=300}
autoplot(web_train) +
  theme_classic() +
  theme(text = element_text(family = 'serif')) +
  labs(y = 'Number of Page Loads',
       title = 'Daily Page Loads')
```

```{r, dpi=300}
autoplot(diff(web_train, lag =7)) +
  theme_classic() +
  theme(text = element_text(family = 'serif')) +
  labs(y = 'Number of Page Loads',
       title = 'Daily Page Loads Seasonally Differenced Once')
```

```{r}
web_1 = diff(web_train, lag =7)
```

```{r}
adf.test(web_1)
```

```{r}
summary(ur.kpss(web_1))
```

## Identify Candidate Models 

```{r, dpi=300}
par(mfrow = c(1,2), family = 'serif')
acf(web_ts)
pacf(web_ts)
```

```{r, dpi=300}
par(mfrow = c(1,2), family = 'serif')
acf(web_1)
pacf(web_1)
dev.off()
```

## Fit Candidate Model and Check Diagnostics 

```{r}
model1 = Arima(web_train,
                    order = c(5, 0, 6),
                    seasonal = list(order = c(1, 1, 1), period = 7))
model1
```

```{r, dpi=300, fig.height=7}
tsdiag(model1)
```

```{r, dpi=300}
checkresiduals(model1)
```

```{r, dpi=300}
# Forecast the next 30 days
my_forecast <- forecast(model1, h = pred_num)

# Plot it!
par(family = 'serif')
plot(my_forecast)
```

```{r}
page_loads_accuracy <- as.data.frame(accuracy(my_forecast, web_test)) %>%
  mutate(variable = 'page_loads')
accuracy(my_forecast, web_test)
```

## Daily Unique Visits 

```{r isolate unique visits}
# unique variable 
unique <- ts(select(traffic, Unique.Visits),
             frequency = 7)
#split into training and testing sets 
unique.train <- head(unique, train_num)
unique.test <- tail(unique, pred_num)
```

```{r, dpi=300}
autoplot(unique.train) +
  theme_classic() +
  labs(y = 'Number of Unique Visits',
       title = 'Daily Unique Website Visits') +
  theme(text = element_text(family = 'serif'))
```

The time series plot appears to show seasonal fluctuations, which means we need to take the seasonal difference before fitting the model. 

```{r, dpi=300}
par(mfrow = c(1,2), family = 'serif')
acf(unique.train)
pacf(unique.train)
```

Fluctuations in the ACF plot also suggest a seasonal component based on the pattern of significant autocorrelation coefficients spiking at each seasonal lag. 


```{r}
adf.test(unique.train)
```

According to the ADF test, the data are stationary; however, this test is not sensitive to seasonal fluctuations. This indicates there may be no long-term trend in the data. 


```{r}
summary(ur.kpss(unique.train))
```

The KPSS test suggests the data are non-stationary. This may be a result of a long-term trend that is not yet detected or a result of the seasonal component influencing the stationary measure. We will take one seasonal difference and re-evaluate the data. 

```{r, dpi=300}
unique.sdiff.1 <- diff(unique.train, lag = 7)
par(mfrow = c(1,2), family = 'serif')
acf(unique.sdiff.1) 
pacf(unique.sdiff.1)
```

After taking the seasonal difference, the ACF appears to show a sinous shape, and the PACF appears to be tailing off with spikes at the seasonal lags. The PACF suggests a seasonal moving average component to the model as well as a regular moving average component because there are significant partial correlation coefficients surrounding the seasonal lags. 

```{r, dpi=300}
autoplot(unique.sdiff.1) +
  theme_classic() +
  theme(text = element_text(family = 'serif'))
```

The time series plot appears to show less seasonal fluctuation now and appears to be stationary. 

```{r}
adf.test(unique.sdiff.1)
```

The ADF test suggests the data are stationary.

```{r}
summary(ur.kpss(unique.sdiff.1))
```

The KPSS test also suggests the data are stationary. We will now identify some candidate models to fit to this variable. 

### identify candidate models 

We will use the auto.arima() function to identify one candidate model. 

```{r}
unique.auto <- auto.arima(unique.train)
unique.auto
```

The auto.arima() function identified a SARIMA(1,0,3) x (0,1,2) model. 

Based on the ACF and PACF plots, we will fit the following three candidate models: 
* SARIMA(5,0,1) x (2,1,2)
* SARIMA(1,0,2) x (2,1,4)
* SARIMA(2,0,2) x (0,1,1) - from auto.arima() 

### Fit Candidate Models and Evaluate Diagnostics 

SARIMA(5,0,1) x (2,1,2)

```{r, dpi=300, fig.height=7}
unique.501.212 <- Arima(unique.train, order = c(5,0,1), seasonal = c(2,1,2))
tsdiag(unique.501.212)
```

Based on the model diagnostics, this model is not appropriate for the data. The Ljung-Box test shows many significant p-values, indicating that the residuals are not independent. 

SARIMA(1,0,2) x (2,1,4)

```{r, dpi=300, fig.height=7}
unique.102.214 <- Arima(unique.train, order = c(1,0,2), seasonal = c(2,1,4))
tsdiag(unique.102.214)
```

The model diagnostics of this model also indicate it is not an appropriate fit to the data based on the significant Ljung-Box p-values. 

SARIMA(1,0,3) x (0,1,2) 

```{r, dpi=300, fig.heigh=7}
tsdiag(unique.auto)
```

The model diagnostics for the auto.arima() model also show that it is not an appropriate fit for the data based on the Ljung-Box p-values. 

We will increase the order to see if the model becomes more appropriate for the data. 

```{r, dpi=300, fig.height=7}
unique.506.213 <- Arima(unique.train, 
                        order = c(5,0,6),
                        seasonal = c(2,1,3))
tsdiag(unique.506.213)
```

```{r, dpi=300}
checkresiduals(unique.506.213)
```

Based on the model diagnostics, this is an appropriate model for the data. 

Now, we will predict the next 217 observations. 

```{r}
unique.forecast <- forecast(unique.506.213, h = pred_num)
```

```{r, dpi=300}
autoplot(unique) +
  autolayer(unique.forecast, alpha = .5) +
  theme_classic() +
  theme(text = element_text(family = 'serif'))
```

The prediction interval covers the actual data, but the point predictions do not follow the fluctuations in the data. 

```{r}
unique.accuracy <- as.data.frame(accuracy(unique.forecast, unique.test)) %>%
  mutate(variable = 'unique_visits')
accuracy(unique.forecast, unique.test)
```

## Returning Visits 

```{r}
return <- ts(select(traffic, Returning.Visits),
             freq = 7) 
#split into training and testing sets 
return.train <- head(return, train_num)
return.test <- tail(return, pred_num)
```

```{r, dpi=300}
autoplot(return.train) +
  theme_classic() +
  theme(text = element_text(family = 'serif'))
```

The time series plot appears to show seasonal fluctuations but no long-term trend. We will test whether this data is already stationary.

```{r}
adf.test(return.train)
```

The ADF test indicates the data are stationary.

```{r}
summary(ur.kpss(return.train))
```

The KPSS test indicates the data are not stationary. This means some differencing may be required. We will start by taking one seasonal difference. 

```{r, dpi=300}
return.sdiff.1 <- diff(return, lag = 7)
autoplot(return.sdiff.1) +
  theme_classic() +
  theme(text = element_text(family = 'serif'))
```

Taking the seasonal difference appears to remove the cyclical fluctuations from the data, so it looks more stationary. 

## Identify Candidate Models

```{r, dpi=300}
par(mfrow = c(1,2), family = 'serif')
acf(return.sdiff.1)
pacf(return.sdiff.1)
```

The ACF and PACF plots appear to tail off, indicating there may be both an AR and MA component. 

```{r}
adf.test(return.sdiff.1)
```

The ADF test indicates the seasonally differenced data are stationary.

```{r}
summary(ur.kpss(return.sdiff.1))
```

The KPSS test also indicates that the seasonally differenced data are stationary. We will now identify some candidate models to fit to the data. Since we have taken the seasonal difference, we may need to identify seasonal AR and MA components as well as regular components. 

Based on the ACF and PACF plots, candidate models we will fit to the data are 
* SARIMA(5,0,3) x (3,1,4)
* SARIMA(6,0,5) x (4,1,3) 

We will also evaluate the model chosen by the auto.arima() function. 

## Fit Candidate Models and Run Diagnostics 

```{r, dpi=300, fig.height=7}
return.auto <- auto.arima(return.train)
tsdiag(return.auto)
```

The model chosen by auto.arima() does not appear to be an appropriate fit for the data based on diagnostic criteria.  

SARIMA(5,0,3) x (3,1,4)
```{r, dpi=300, fig.height=7}
return.503.314 <- Arima(return.train, order = c(5,0,3), seasonal = c(3,1,4))
tsdiag(return.503.314)
```

Based on the diagnostic criteria, this model is not an appropriate fit to the data. There are significant p-values in the Ljung-Box test, indicating the residuals are not independent. 

SARIMA(6,0,5) x (4,1,3)
```{r, dpi=300, fig.height=7}
return.605.413 <- Arima(return.train, order = c(6,0,5), seasonal = c(4,1,3))
tsdiag(return.605.413)
```

The diagnostic criteria indicate that this model is an appropriate fit to the data. We will also check for normal distribution of residuals. 

```{r, dpi=300}
checkresiduals(return.605.413)
```

The residuals appear to follow a normal distribution, which provides further support for the model. This is the most appropriate fit for the data. 

Now, we will predict the next 217 observations using the chosen model. 

```{r, dpi=300}
return.forecast <- forecast(return.605.413, h = pred_num)
autoplot(return) +
  autolayer(return.forecast, alpha = .5) +
  theme_classic() +
  theme(text = element_text(family = 'serif'))
```

The forecasted values follow the fluctuations of the data and the real data fall within the prediction intervals. 

```{r}
return.accuracy <- as.data.frame(accuracy(return.forecast, return.test)) %>%
  mutate(variable = 'return_visits')
accuracy(return.forecast, return.test)
```

# Multivariate Time Series Model 

```{r, dpi=300}
multi <- ts(select(traffic, Page.Loads, Unique.Visits, Returning.Visits))
autoplot(multi, facet = TRUE) +
  theme_minimal() +
  theme(text = element_text(family = 'serif'))
```

Like the univariate analyses, the variables do not look stationary, so we will take one seasonal difference of each variable. 

```{r, dpi=300}
multi_sdiff_1 <- diff(multi, lag = 7)
autoplot(multi_sdiff_1, facet = TRUE) +
  theme_minimal() +
  theme(text = element_text(family = 'serif'))
```

The seasonally difference data appears more stationary than the raw data. 

```{r}
multi_train <- head(multi_sdiff_1, train_num)
multi_test <- tail(multi_sdiff_1, pred_num)
```

## Identify and Fit Candidate Models 

We will examine the cross-correlations of the data to identify candidate models. 

```{r, dpi=300}
ccm(multi_train)
```

The cross-correlation matrix shows significant cross-correlations through lag 6. 

We will start with a VARMA(1,0) model. 

```{r}
multi_fit.1 <- VARMA(multi_train, p = 1, q = 0)
ccm(multi_fit.1$residuals)
enter
```

There are significant cross-correlations among the residuals, so we need to increase the order of the model. 

```{r}
multi_fit.2 <- VARMA(multi_train, p = 2, q = 0)
ccm(multi_fit.2$residuals)
enter
```

Significant cross-correlations among residuals suggest this model is not an appropriate fit for the data. 

```{r}
multi_fit.3 <- VARMA(multi_train, p = 3, q = 0)
ccm(multi_fit.3$residuals)
enter
```

There are fewer significant cross-correlations, but this model is still not appropriate for the data. 

```{r}
multi_fit.4 <- VARMA(multi_train, p = 4, q = 0)
ccm(multi_fit.4$residuals)
enter
```

There are still significant cross-correlations so this model is not a good fit. We will continue to increase the AR order to find a more appropriate fit for the data. 

```{r}
multi_fit.5 <- VARMA(multi_train, p = 5, q = 0)
ccm(multi_fit.5$residuals)
enter
```

```{r}
multi_fit.6 <- VARMA(multi_train, p = 6, q = 0)
ccm(multi_fit.6$residuals)
enter
```

```{r}
multi_fit.6.1 <- VARMA(multi_train, p = 6, q = 1)
ccm(multi_fit.6.1$residuals)
enter
```

```{r}
multi_fit.6.2 <- VARMA(multi_train, p = 6, q = 2)
ccm(multi_fit.6.2$residuals)
enter
```

```{r}
multi_fit.7 <- VARMA(multi_train, p = 7, q = 0)
ccm(multi_fit.7$residuals)
enter
```

```{r}
multi_fit.8 <- VARMA(multi_train, p = 8, q = 0)
ccm(multi_fit.8$residuals)
enter
```

```{r}
multi_fit.9 <- VARMA(multi_train, p = 9, q = 0)
ccm(multi_fit.9$residuals)
enter
```

```{r}
multi_fit.10 <- VARMA(multi_train, p = 9, q = 1)
ccm(multi_fit.10$residuals)
enter
```

```{r}
multi_fit.11 <- VARMA(multi_train, p = 9, q = 2)
ccm(multi_fit.11$residuals)
enter
```

```{r}
multi_fit.12 <- VARMA(multi_train, p = 10, q = 0)
ccm(multi_fit.12$residuals)
enter
```

```{r}
multi_fit.13 <- VARMA(multi_train, p = 11, q = 0)
ccm(multi_fit.13$residuals)
enter
```

```{r}
multi_fit.14 <- VARMA(multi_train, p = 12, q = 0)
ccm(multi_fit.14$residuals)
enter
```

There are still two significant cross-correlations at lags 7 and 12. To improve the model fit, we will add a moving average component to address the cross-correlation at lag 7. 

```{r}
multi_fit.15 <- VARMA(multi_train, p = 12, q = 7)
ccm(multi_fit.15$residuals)
enter
```

Adding a moving average component MA(7) improved the fit of the model, and there are no longer any significant cross-correlations. 

We will now use the identified model to predict the next 217 observations of each variable individually. 

```{r}
# page loads 
page_loads_pred <- VARMApred(multi_fit.15, h = pred_num)$pred[,1]
page_loads_sd <- VARMApred(multi_fit.15, h = pred_num)$se.err[,1]
page_loads_forecast <- cumsum(c(multi[1,1], multi_train[,1], page_loads_pred)) 
```

```{r, dpi=300}
par(family = 'serif')
plot(page_loads_forecast[1:1950], xlim = c(0,2200), type = 'l', lwd = 2,
     main = 'Forecasted Page Loads', xlab = 'Time', ylab = 'Number of Page Loads')
lines(page_loads_forecast[1950:2168], x=seq(1950, 2168, 1), type = 'l', col = 'red', lty = 2, lwd = 2)
lines(page_loads_forecast[1950:2168]+(2*page_loads_sd), x=seq(1950, 2168, 1), col = 'blue', type = 'l', lwd = 2, lty = 3)
lines(page_loads_forecast[1950:2168]-(2*page_loads_sd), x=seq(1950, 2168, 1), col = 'blue', type = 'l', lwd = 2, lty = 3)
legend('bottomright', legend = c('Page Loads Data', 'Forecasted Page Loads', '95% Confidence Interval'), col = c(1,2,'blue'), lty = c(1,2,3), lwd = c(2,2,2))
```

```{r}
unique_visits_pred <- VARMApred(multi_fit.15, h = pred_num)$pred[,2]
unique_visits_sd <- VARMApred(multi_fit.15, h = pred_num)$se.err[,2]
unique_visits_forecast <- cumsum(c(multi[1,2], multi_train[,2], unique_visits_pred))
```

```{r, dpi=300}
par(family = 'serif')
plot(unique_visits_forecast[1:1950], xlim = c(0,2200), type = 'l', lwd = 2,
     main = 'Forecasted Unique Visits', xlab = 'Time', ylab = 'Number of Unique Visits')
lines(unique_visits_forecast[1950:2168], x=seq(1950, 2168, 1), type = 'l', col = 'red', lty = 2, lwd = 2)
lines(unique_visits_forecast[1950:2168]+(2*unique_visits_sd), x=seq(1950, 2168, 1), col = 'blue', type = 'l', lwd = 2, lty = 3)
lines(unique_visits_forecast[1950:2168]-(2*unique_visits_sd), x=seq(1950, 2168, 1), col = 'blue', type = 'l', lwd = 2, lty = 3)
legend('bottomright', legend = c('Unique Visits Data', 'Forecasted Unique Visits', '95% Confidence Interval'), col = c(1,2,'blue'), lty = c(1,2,3), lwd = c(2,2,2))
```

```{r}
returning_visits_pred <- VARMApred(multi_fit.15, h = pred_num)$pred[,3]
returning_visits_sd <- VARMApred(multi_fit.15, h = pred_num)$se.err[,3]
returning_visits_forecast <- cumsum(c(multi[1,3], multi_train[,3], returning_visits_pred))
```

```{r, dpi=300}
par(family = 'serif')
plot(returning_visits_forecast[1:1950], xlim = c(0,2200), type = 'l', lwd = 2,
     main = 'Forecasted Returning Visits', xlab = 'Time', ylab = 'Number of Returning Visits')
lines(returning_visits_forecast[1950:2168], x=seq(1950, 2168, 1), type = 'l', col = 'red', lty = 2, lwd = 2)
lines(returning_visits_forecast[1950:2168]+(2*returning_visits_sd), x=seq(1950, 2168, 1), col = 'blue', type = 'l', lwd = 2, lty = 3)
lines(returning_visits_forecast[1950:2168]-(2*returning_visits_sd), x=seq(1950, 2168, 1), col = 'blue', type = 'l', lwd = 2, lty = 3)
legend('bottomright', legend = c('Returning Visits Data', 'Forecasted Returning Visits', '95% Confidence Interval'), col = c(1,2,'blue'), lty = c(1,2,3), lwd = c(2,2,2))
```

```{r}
multi_pageloads_accuracy <- as.data.frame(accuracy(page_loads_forecast, web_test)) %>%
  mutate(variable = 'VARMA_page_loads')
accuracy(page_loads_forecast, web_test)
```

```{r}
multi_uniquevisits_accuracy <- as.data.frame(accuracy(unique_visits_forecast, unique.test)) %>%
  mutate(variable = 'VARMA_unique_visits')
```

```{r}
multi_returnvisits_accuracy <- as.data.frame(accuracy(returning_visits_forecast, return.test)) %>%
  mutate(variable = 'VARMA_returning_visits')
```

```{r}
multi_accuracy <- rbind(multi_pageloads_accuracy,
                        multi_uniquevisits_accuracy,
                        multi_returnvisits_accuracy)
flextable(multi_accuracy %>% rownames_to_column('Set'))
```


# Dynamic Regression Model 


# Comparing Model Accuracies 

```{r}
univ_accuracy <- rbind(page_loads_accuracy, 
                  unique.accuracy,
                  return.accuracy)
accuracy <- rbind(univ_accuracy[,-6],
                  multi_pageloads_accuracy,
                  multi_uniquevisits_accuracy,
                   multi_returnvisits_accuracy)
flextable(accuracy %>% rownames_to_column('Set'))
```

Based on the accuracy metrics, the univariate analyses appear to produce more accurate forecast predictions than the multivariate VARMA model. The VARMA model included a very large number of parameters, which may have caused overfitting to the training data, making the predictions less accurate. 