---
title: "Final Project"
author: "AFN"
date: "2026-04-17"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(dplyr)
library(tseries)
library(fpp2)
library(ggplot2)
library(urca)
library(TSA)
```

```{r read in data}
traffic <- read.csv('Data/daily-website-visitors.csv')
```

# Univariate Unique Visits 

```{r isolate unique visits}
unique <- ts(select(traffic, Unique.Visits), 
             freq = 365,
             start = c(2014, 257))
```

```{r}
autoplot(unique) 
```

Check if variable is stationary 

```{r}
par(mfrow = c(1,2), family = 'serif')
acf(unique)
pacf(unique)
```


```{r}
adf.test(unique)
```

ADF test says data are stationary.


```{r}
summary(ur.kpss(unique))
```

KPSS says data are not stationary.

ACF plot suggests there are seasonal fluctuations in the data. 

We will take the seasonal difference of the data before checking for stationarity. 

```{r}
unique.diff.1 <- diff(unique)
autoplot(unique.diff.1)
```

```{r}
par(mfrow = c(1,2))
acf(unique.diff.1)
pacf(unique.diff.1)
```

ACF plot is kind of crazy 

```{r}
adf.test(unique.diff.1)
```

ADF says data are stationary. 

```{r}
summary(ur.kpss(unique.diff.1))
```

KPSS also says data are stationary.

```{r}
nsdiffs(unique)
```

Function says number of seasonal differences to take is 0. 

```{r}
unique.sdiff.1 <- diff(unique, lag = 7)
par(mfrow = c(1,2), family = 'serif')
acf(unique.sdiff.1) 
pacf(unique.sdiff.1)
```

```{r}
autoplot(unique.sdiff.1)
```


```{r}
adf.test(unique.sdiff.1)
```

ADF says data are stationary.

```{r}
summary(ur.kpss(unique.sdiff.1))
```

KPSS also says data are stationary.

```{r}
unique.sdiff.2 <- diff(unique.sdiff.1, lag = 7)
par(mfrow = c(1,2), family = 'serif')
acf(unique.sdiff.2) 
pacf(unique.sdiff.2)
```

```{r}
autoplot(unique.sdiff.2)
```

```{r}
adf.test(unique.sdiff.2)
```

ADF says data are stationary.

```{r}
summary(ur.kpss(unique.sdiff.2))
```

KPSS says data are stationary.

## identify candidate models 


```{r}
eacf(unique.sdiff.2)
```

Possible models based on EACF are AR(1), MA(1); or AR(0), MA(2) 

Seasonal components: MA(2), AR(2)

```{r}
auto.arima(unique)
```

Auto ARIMA does not include a seasonal component or regular differencing. 

CANDIDATE MODELS 
* SARIMA(1,0,1) x (2,2,2)
* SARIMA(0,0,2) x (2,2,2)



